### 2023-10-19 about.qmd
Update

---
title: "Data Story"
---

<br>

### Investigating Singapore's Resale Flat Prices from 1990 to 2023

##### Is public housing in Singapore affordable for the average Singaporean?

<br> This is where my introduction will be. This is just a placeholder for potential data sets!

```{r, eval=TRUE,echo=FALSE}
library(tidyverse)
starwars
```