### 2023-10-27 about.qmd
Project Update

---
title: "Data Story"
---

## Is Public Housing in Singapore Affordable for the Average Singaporean?

<br>

### Investigating Singapore's Resale Flat Prices from 1990 to 2023

In Singapore, housing affordability has been an increasing worry for future Singaporeans. However, in 2012, the then Deputy Prime Minister (DPM) Tharman Shanmugaratnam - stating household income of SGD \$1000 was sufficient to own a home, sparked discussions among netizen online (Tan, 2012). While Mr Tharman argued the housing grants could provide support for low-income household, many Singaporeans expressed concern about the long-term viability of such claims, especially in the face of the constant rising cost of living and property prices.

<br> This is where my introduction will be. This is just a placeholder for potential data sets!

```{r, eval=TRUE,echo=FALSE}
library(tidyverse)
starwars
```