### 2023-11-23 about.qmd
Project Update

---
title: "My Experiences"
---

![](src/images/aboutme.png)

## My Work and Internship Experience - [My LinkedIn Profile](https://www.linkedin.com/in/marzuki-nooranas-b5947b17a/)

**- Specialist**\
*Apple South Asia* <br> Aug 2023 - Present · 4 mos

**- Galaxy Specialist · Part-time**\
*Samsung Electronics Singapore* <br>Feb 2018 - Jan 2023 · 5 yrs

**- Business Information Technology Interest Group - Honorary Secretary**\
*Temasek Polytechnic* <br>Apr 2019 - Apr 2021 · 2 yrs 1 mo

**- Application Development Intern**\
*Accenture · Internship* <br>Apr 2020 - Aug 2020 · 5 mos

**- Marketing Research Intern**\
*Shopee Jakarta, Indonesia· Internship* <br>Mar 2020 - Apr 2020 · 2 mos

**- Instructor**\
*Singapore Civil Defence Force · Full-time* <br>Feb 2016 - Nov 2017 · 1 yr 10 mos

**- Attraction Host Lead**\
*Sentosa Development Corporation · Part-time* <br>Oct 2013 - Nov 2015 · 2 yrs 2 mos