### 2023-11-24 about.qmd
Project Update

---
title: "My Experiences"
---

![](src/images/aboutme.png)

## Education

**National University of Singapore**\
*Bachelor of Social Sciences \[Hons.\], Communications New Media / Sociology, Social Sciences* <br> Aug 2022 - Present

**Temasek Polytechnic**\
*Diploma with Merit, Business Information Technology* <br> Apr 2018 - Apr 2021

**ITE College Central**\
*Higher Nitec in Business Event Management* <br> Apr 2013 - Apr 2015

## Work & Internship - [LinkedIn Page](https://www.linkedin.com/in/marzuki-nooranas-b5947b17a/)

**Product Specialist**\
*Apple South Asia* <br> Aug 2023 - Present · 4 mos

**Galaxy Specialist · Part-time**\
*Samsung Electronics Singapore* <br>Feb 2018 - Jan 2023 · 5 yrs

**Business Information Technology Interest Group - Honorary Secretary**\
*Temasek Polytechnic* <br>Apr 2019 - Apr 2021 · 2 yrs 1 mo

**Application Development Intern**\
*Accenture · Internship* <br>Apr 2020 - Aug 2020 · 5 mos

**Marketing Research Intern**\
*Shopee Jakarta, Indonesia· Internship* <br>Mar 2020 - Apr 2020 · 2 mos

**Instructor**\
*Singapore Civil Defence Force · Full-time* <br>Feb 2016 - Nov 2017 · 1 yr 10 mos

**Attraction Host Lead**\
*Sentosa Development Corporation · Part-time* <br>Oct 2013 - Nov 2015 · 2 yrs 2 mos