### 2025-01-28 index.qmd
Update

---
title: "Well Hello There!"
---

```{r, eval=TRUE,echo=FALSE}

```

![](src/images/namelogo.png){fig-align="left" width="800"}

![](src/images/me2.jpg)

## My Hello to You \~\~

Hello there! My name is Marzuki Nooranas \[[Resume](https://raw.githubusercontent.com/mrzknrns/mrzknrns.github.io/c446afc4b2d7902cfc76a22458cbc7789d1a09d7/MARZUKI_SANIP_RESUME.pdf)\] - or *Marz* for short! I am currently in my 3rd year, majoring in Sociology and Interactive Media Development!

I have always been fascinated by how society functions and its intricate connections, my passion for understanding the world around us, is driven by a desire to foster meaningful connections within all communities!

## My overland journey from Singapore to Vietnam!

To demonstrate this, I embark of frequent backpacking trips across Southeast Asia, and actively engage myself in the cultures and learn about their customs.

Such experiences provide me with firsthand insights into how different societies interact, and not only broadening my horizons, but also fostering empathy across different communities. Here's a little snippet -

```{=html}
<style>
body, html {
  margin: 0;
  padding: 0;
  overflow-x: hidden; 
}

.tiktok-container {
  display: flex;
  flex-wrap: wrap;
  justify-content: center;
  width: 105%; 
}

.tiktok-embed {
  box-sizing: border-box;
  max-width: 100%;
  flex: 1 1 300px;
  margin: 0 auto; 
}

@media screen and (max-width: 600px) {
  .tiktok-embed {
    min-width: auto;
  }
}
</style>
```
::: tiktok-container
<blockquote class="tiktok-embed" cite="https://www.tiktok.com/@mentaikofries_/video/7181240142063996161" data-video-id="7181240142063996161" style="max-width: 605px;min-width: 325px;">

<section>

<a target="_blank" title="@mentaikofries_" href="https://www.tiktok.com/@mentaikofries_?refer=embed">@mentaikofries_</a>

<p>overland travel for the vibesss</p>

<a target="_blank" title="♬ original sound - 𝚁𝚊𝚐𝚑𝚊𝚍💗" href="https://www.tiktok.com/music/original-sound-7178872451673459457?refer=embed">♬ original sound - 𝚁𝚊𝚐𝚑𝚊𝚍💗</a>

</section>

</blockquote>

<blockquote class="tiktok-embed" cite="https://www.tiktok.com/@mentaikofries_/video/7383243879715196168" data-video-id="7383243879715196168" style="max-width: 605px;min-width: 325px;">

<section>

<a target="_blank" title="@mentaikofries_" href="https://www.tiktok.com/@mentaikofries_?refer=embed">@mentaikofries_</a>

<p>errr i drank about 9 cups of pearl milk tea in 4 days 3 nights 🇹🇼🥴🧋🇹🇼</p>

<a target="_blank" title="♬ fish in the pool・花屋敷 - ヘクとパスカル" href="https://www.tiktok.com/music/fish-in-the-pool-花屋敷-6910585281334740994?refer=embed">♬ fish in the pool・花屋敷 - ヘクとパスカル</a>

</section>

</blockquote>

```{=html}
<script async src="https://www.tiktok.com/embed.js"></script>
```
:::

## Contact Details

For any opportunities:

Email - marzukinooranas\@gmail.com

Linkedin - [Marzuki Nooranas](https://www.linkedin.com/in/marzuki-nooranas-b5947b17a/)

Resume - [Download](https://raw.githubusercontent.com/mrzknrns/mrzknrns.github.io/c446afc4b2d7902cfc76a22458cbc7789d1a09d7/MARZUKI_SANIP_RESUME.pdf)

```         
```