### 2025-01-28 interests.qmd
Update

---
title: "My Past Work"
---

```{r, eval=TRUE,echo=FALSE}
```

<br> I've dabbled in UX/UI design and video editing, mostly picking up these skills on my own because I'm really into it. I love creating cool user experiences and videos that pop, and I'm excited to keep leveling up and bringing more creativity to my work!

# UX/UI Prototyping and Design Work

## Interaction Design

### (1) SlideCraft - AI Powered Design Tool

SlideCraft is an AI-powered slide design tool created to simplify the presentation-making process for students and educators. I developed detailed lo-fi and hi-fi prototypes, focusing on features like design consistency, personalized recommendations, and efficiency.

This project emphasized enabling users to focus on content creation while saving time with AI-driven slide suggestions.

<iframe style="border: 1px solid rgba(0, 0, 0, 0.1);" width="800" height="450" src="https://embed.figma.com/proto/4o0fMCNtIOXtFuupnMgmlU/Slidecraft-Hi-Fi-Prototype-Flow?page-id=0%3A1&amp;node-id=1-31&amp;starting-point-node-id=1%3A31&amp;embed-host=share">

</iframe>

### (2) GPTDashboard - Desktop View

The ChatGPT Dashboard was designed to help social media content creators streamline their workflow by integrating AI into content generation and tracking. I worked on both lo-fi and hi-fi prototypes, ensuring a smooth user experience across desktop and mobile views.

This project focused on addressing key pain points like burnout and content optimization while providing real-time feedback metrics to help creators stay on top of trends​

<iframe style="border: 1px solid rgba(0, 0, 0, 0.1);" width="800" height="450" 
src="https://embed.figma.com/proto/LZcst4QtAgGJgik4eHhaRb/ChatGPT-Dashboard-Hi-Fi-Desktop-View-Prototype-Link?page-id=0%3A1&node-id=1-228&starting-point-node-id=1%3A228&embed-host=share" 
allowfullscreen>

</iframe>

### (3) SaveWatt

SaveWatt is a digital platform aimed at helping homeowners make smarter, energy-efficient appliance choices amidst rising electricity tariffs. I contributed to the design of a user-friendly interface that consolidates appliance comparison tools, offering personalized recommendations and energy-saving insights.

This project tackled the challenges of decision fatigue and financial strain by providing actionable guidance in a visually engaging way.

Full PortFolio: <https://savewatt.framer.website>

![](src/images/SaveWatt.png)

<iframe style="border: 1px solid rgba(0, 0, 0, 0.1);" width="800" height="450" 
src="https://embed.figma.com/proto/33D1uDg1g786wbtU9np2gq/4_G3_Prototype?page-id=3%3A646&node-id=89-2546&starting-point-node-id=89%3A2546&embed-host=share" 
allowfullscreen>

</iframe>

## TeachSG

For my TeachSG attachment, I created this motivational card specifically for primary school students who were struggling with math. The card was designed to inspire confidence and encourage them as they prepared for their PSLE exams.

Using bright colors, cheerful designs, and a personalized touch with mentor memojis, I aimed to make the card relatable and uplifting, helping the students feel supported and motivated to achieve their best.

::: {style="position: relative; width: 100%; height: 0; padding-top: 56.2225%; padding-bottom: 0; box-shadow: 0 2px 8px 0 rgba(63,69,81,0.16); margin-top: 1.6em; margin-bottom: 0.9em; overflow: hidden; border-radius: 8px; will-change: transform;"}
<iframe loading="lazy" style="position: absolute; width: 100%; height: 100%; top: 0; left: 0; border: none; padding: 0;margin: 0;" src="https://www.canva.com/design/DAGRwrKGhxo/DFyJSp72gAFRLvh1wU3bwA/view?embed" allowfullscreen="allowfullscreen" allow="fullscreen">

</iframe>
:::

<a href="https://www.canva.com/design/DAGRwrKGhxo/DFyJSp72gAFRLvh1wU3bwA/view?utm_content=DAGRwrKGhxo&utm_campaign=designshare&utm_medium=embeds&utm_source=link" target="_blank" rel="noopener"></a>

## Artificial Intelligence and Society

In this module, we looked at the benefits and potential downsides of relying on AI. My group focused on leveraging on AI technology to aid in bereavement process. I designed a brochure for a Virtual Memory Garden where users will transported into a virtual space together with their loved ones - as an avatar!

::: {style="position: relative; width: 100%; height: 0; padding-top: 56.2225%; padding-bottom: 0; box-shadow: 0 2px 8px 0 rgba(63,69,81,0.16); margin-top: 1.6em; margin-bottom: 0.9em; overflow: hidden; border-radius: 8px; will-change: transform;"}
<iframe loading="lazy" style="position: absolute; width: 100%; height: 100%; top: 0; left: 0; border: none; padding: 0;margin: 0;" src="https://www.canva.com/design/DAF0lbgxhPk/jKGt5Uqtio2trLaLP-h9tw/view?embed" allowfullscreen="allowfullscreen" allow="fullscreen">

</iframe>
:::

```         
```

```         
```