### 2025-08-05 index.qmd
Resume Update

---
title: "Well Hello There!"
---

```{r, eval=TRUE,echo=FALSE}

```

![](src/images/namelogo.png){fig-align="left" width="800"}

![](src/images/me2.jpg)

## My Hello to You \~\~

Hello there! My name is Marzuki Nooranas or *Marz* for short! I am currently in my final year, majoring in Sociology and Interactive Media Development!

I have always been fascinated by how society functions and its intricate connections, my passion for understanding the world around us, is driven by a desire to foster meaningful connections within all communities!

## My overland journey from Singapore to Vietnam!

To demonstrate this, I embark on frequent backpacking trips across Southeast Asia and actively immerse myself in local cultures to learn about their customs. These experiences provide me with firsthand insights into how different societies interact, not only broadening my horizons but also fostering empathy across diverse communities. Here's a little snippet!

```{=html}
<style>
  body, html {
    margin: 0;
    padding: 0;
    overflow-x: hidden;
  }

  .tiktok-container {
    display: flex;
    flex-wrap: wrap;
    justify-content: space-between; /* Distribute space between items */
    gap: 20px; /* Add spacing between items */
    width: 100%;
    max-width: 1200px; /* Limit maximum width for larger screens */
    margin: 0 auto; /* Center the container */
  }

  .tiktok-embed {
    box-sizing: border-box;
    flex: 1 1 calc(50% - 10px); /* Two items per row with spacing */
    max-width: calc(50% - 10px); /* Two items per row with spacing */
    min-width: 300px; /* Minimum width for each embed */
  }

  @media screen and (max-width: 600px) {
    .tiktok-embed {
      flex: 1 1 100%; /* Full width on mobile */
      max-width: 100%; /* Full width on mobile */
    }
  }
</style>
```
::: tiktok-container
<blockquote class="tiktok-embed" cite="https://www.tiktok.com/@mentaikofries_/video/7181240142063996161" data-video-id="7181240142063996161">

<section>

<a target="_blank" title="@mentaikofries_" href="https://www.tiktok.com/@mentaikofries_?refer=embed">@mentaikofries_</a>

<p>overland travel for the vibesss</p>

<a target="_blank" title="♬ original sound - 𝚁𝚊𝚐𝚑𝚊𝚍💗" href="https://www.tiktok.com/music/original-sound-7178872451673459457?refer=embed">♬ original sound - 𝚁𝚊𝚐𝚑𝚊𝚍💗</a>

</section>

</blockquote>

<blockquote class="tiktok-embed" cite="https://www.tiktok.com/@mentaikofries_/video/7383243879715196168" data-video-id="7383243879715196168">

<section>

<a target="_blank" title="@mentaikofries_" href="https://www.tiktok.com/@mentaikofries_?refer=embed">@mentaikofries_</a>

<p>errr i drank about 9 cups of pearl milk tea in 4 days 3 nights 🇹🇼🥴🧋🇹🇼</p>

<a target="_blank" title="♬ fish in the pool・花屋敷 - ヘクとパスカル" href="https://www.tiktok.com/music/fish-in-the-pool-花屋敷-6910585281334740994?refer=embed">♬ fish in the pool・花屋敷 - ヘクとパスカル</a>

</section>

</blockquote>
:::

```{=html}
<script async src="https://www.tiktok.com/embed.js"></script>
```
## Contact Details

For any opportunities:

Email - marzukinooranas\@gmail.com

Linkedin - [Marzuki Nooranas](https://www.linkedin.com/in/marzuki-nooranas-b5947b17a/)

Resume - [Download](https://raw.githubusercontent.com/mrzknrns/mrzknrns.github.io/c446afc4b2d7902cfc76a22458cbc7789d1a09d7/MARZUKI_SANIP_RESUME.pdf)

```         
```